---
title: "assignment8"
author: "Noah Plant"
date: "2024-09-30"
output: pdf_document
---

## Excersise 1
Create a vector of three elements (2,4,6) and name that vector `vec_a`. Create a second vector, `vec_b`, that contains (8,10,12). Add these two vectors together and name the result `vec_c`.

```{r}
vec_a=c(2,4,6)

vec_b=c(8,10,12)

vec_c=vec_a+vec_b

vec_c
```

## Excersise 2
Create a vector, named `vec_d`, that contains only two elements (14,20). Add this vector to `vec_a`. What is the result and what do you think R did (look up the recycling rule using Google)? What is the warning message that R gives you?

```{r}

vec_d=c(14,20)

vec_a+vec_d

```

It looks like in this instance R added up the vectors element wise. And for the elements that C had that D didnt, R just did nothing.

## Excersise 3
Next add 5 to the vector vec_a. What is the result and what did R do? Why doesn't in give you a warning message similar to what you saw in the previous problem?

```{r}
vec_a+5

```

In this instance R treated 5 like the vector (5,5,5), and did not give a warning. This is because R treats the scalar object 5 differently then a vector object.

## Excersise 5
Generate the vector of even numbers $\left\{ 2,4,6,\dots,20\right\}$ 
    a) Using the seq() function and 
    
```{r}

vec_5a=seq(from=2, to=20, by=2)

vec_5a

```
    
    
  b) Using the a:b shortcut and some subsequent algebra. *Hint: Generate the vector 1-10 and then multiple it by 2*.
    
```{r}

vec_5b<-1:10

vec_5b<-vec_5b*2

vec_5b

```


## Excersise 6
Generate a vector of 21 elements that are evenly placed between 0 and 1 using the `seq()` command and name this vector `x`. 

```{r}

x<-seq(0,1,length.out=21)
x

```


## Excersise 8
Generate the vector $\left\{ 2,2,2,2,4,4,4,4,8,8,8,8\right\}$
using the `rep()` command. You might need to check the help file for rep() to see all of the options that rep() will accept. In particular, look at the optional argument `each=`.
  
```{r}

vec_8<-c(2,4,8)

vec_8<-rep(vec_8,each=4)
vec_8

```
  
## Excersise 9
Create and manipulate a data frame.
    a) Create a `data.frame` named `my.trees` that has the following columns:
        + Girth = {8.3, 8.6, 8.8, 10.5, 10.7, 10.8, 11.0}
        + Height= {70, 65, 63, 72, 81, 83, 66}
        + Volume= {10.3, 10.3, 10.2, 16.4, 18.8, 19.7, 15.6}
        
        
        
```{r}

my.trees<-data.frame(
  Girth=c(8.3, 8.6, 8.8, 10.5, 10.7, 10.8, 11.0),
  Height=c(70, 65, 63, 72, 81, 83, 66),
  Volume=c(10.3, 10.3, 10.2, 16.4, 18.8, 19.7, 15.6)
)

my.trees


```


        
  b) Without using `dplyr` functions, extract the third observation (i.e. the third row)
    
```{r}
  
my.trees[3,]

```
    
    
  c) Without using `dplyr` functions, extract the Girth column referring to it by name (don't use whatever order you placed the columns in).
    
```{r}

my.trees[,'Girth']

```
    
  d) Without using `dplyr` functions, print out a data frame of all the observations *except* for the fourth observation. (i.e. Remove the fourth observation/row.)
    
```{r}

my.trees2<-my.trees[-c(4),]

my.trees2

```
    
    
  e) Without using `dplyr` functions, use the `which()` command to create a vector of row indices that have a `girth` greater than 10. Call that vector `index`.
    
```{r}

index<-which(my.trees$Girth>10)



```
    
    
  f) Without using `dplyr` functions, use the `index` vector to create a small data set with just the large girth trees.
    
```{r}

my.trees[index,]

```
    
  g) Without using `dplyr` functions, use the `index` vector to create a small data set with just the small girth trees.
    
```{r}
my.trees[-c(index),]
```
    
    
    
    
## Excersise 12

The following code creates a `data.frame` and then has two different methods for removing the rows with `NA` values in the column `Grade`. Explain the difference between the two.
    ```{r, eval=FALSE}
    df <- data.frame(name= c('Alice','Bob','Charlie','Daniel'),
                     Grade = c(6,8,NA,9))
    
    df[ -which(  is.na(df$Grade) ), ]
    
    
    df[  which( !is.na(df$Grade) ), ]
    ```
    
    
In the first line, the program is finding all the rows without a grade and displaying all the rows except that one.

In the second line the program is finding all the rows with a grade and displaying those rows. 

The difference here is subtle, one line uses R's - feature to basically remove rows, while the second line is using the not operator ! to select for the lines that do have grades.

## Excersise 14
Create and manipulate a list.
    a) Create a list named my.test with elements
        + x = c(4,5,6,7,8,9,10)
        + y = c(34,35,41,40,45,47,51)
        + slope = 2.82
        + p.value = 0.000131
        
```{r}

  x = c(4,5,6,7,8,9,10)
  y = c(34,35,41,40,45,47,51)
  slope = 2.82
  p.value = 0.000131
  
  my.test<-list(x.val=x,y.val=y,slope=slope,p.val=p.value)
  
  str(my.test)

```
    
    
    b) Extract the second element in the list.
    
```{r}

my.test[2]

```
    
    c) Extract the element named `p.value` from the list.
    
```{r}

my.test['p.val']

```
    
