---
title: "assignment5"
author: "Noah Plant"
date: "2024-10-28"
output: pdf_document
---
### Exercise 1 {-}

Convert the following to date or date/time objects.

```{r}

library(tidyverse)
library(lubridate)
library(mosaicData)

```


**a)** September 13, 2010.

```{r}


mdy('September 13, 2010')


```


**b)** Sept 13, 2010.

```{r}

as.Date("Sept 13, 2010",format='%b %d, %Y') 

mdy("Sept 13, 2010")

```


**c)** Sep 13, 2010.

```{r}

mdy("Sep 13, 2010")

```


**d)** S 13, 2010. Comment on the month abbreviation needs.

```{r}

mdy("S 13, 2010")



```

The date must be abbreviated to 3 letters, spelled out completely or represented by a 2 digit number. There is no way to format the month as a single letter as that would cause confusion I imagine.

**e)** 07-Dec-1941.

```{r}

dmy("07-Dec-1941")

```


**f)** 1-5-1998. Comment on why you might be wrong.

```{r}
mdy("1-5-1998")
dmy("1-5-1998")


```

I may be wrong in my date making because I have no way of knowing whether the date is supplied in day month year or year month day.

**g)** 21-5-1998. Comment on why you know you are correct.

```{r}

dmy("21-5-1998")

```

I know I am right because if I try to parse the data in any other way I get an error. Also there is no way this date can be written in mdy because there are only 12 months in a year.

**h)** 2020-May-5 10:30 am

```{r}
ymd_hm('2020-May-5 10:30 am')

```


**i)** 2020-May-5 10:30 am PDT (ex Seattle)

```{r}

ymd_hm('2020-May-5 10:30 am PDT')

ymd_hm('2020-May-5 10:30 am PDT',tz="America/Los_Angeles")

```
The time zone does not parse unless I include the tz="PDT" part of the code. Even then if I do that it does not register. I have to include the olson name which is America/Los_Angeles

**j)** 2020-May-5 10:30 am AST (ex Puerto Rico)


```{r}

ymd_hm("2020-May-5 10:30 am AST",tz="America/Puerto_Rico")

```


### Exercise 3 {-}

Suppose you have arranged for a phone call to be at 3 pm on May 8, 2025 at Arizona time. However, the recipient will be in Auckland, NZ. What time will it be there? 

```{r}
  
arizona<-mdy_hm('May 8, 2025 3:00 pm',tz='US/Arizona')

auckland<-with_tz(arizona,tz="Pacific/Auckland")

auckland

```





### Exercise 5 {-}

It turns out there is some interesting periodicity regarding the number of births on particular days of the year.

**a)** Using the `mosaicData` package, load the data set `Births78` which records the number of children born on each day in the United States in 1978. Because this problem is intended to show how to calculate the information using the `date`, remove all the columns *except* `date` and `births`.

```{r}
data<-mosaicData::Births78

data.2<-select(data,date,births)

head(data.2)

```


**b)** Graph the number of `births` vs the `date` with date on the x-axis. What stands out to you? Why do you think we have this trend?

```{r}

library(ggplot2)

ggplot(data=data.2, aes(x=date,y=births) )+
  geom_point()

```


**c)** To test your assumption, we need to figure out the what day of the week each observation is. Use `dplyr::mutate` to add a new column named `dow` that is the day of the week (Monday, Tuesday, etc). This calculation will involve some function in the `lubridate` package and the `date` column.

```{r}

data.2<- data.2 %>% mutate(dow=(wday(date,label=TRUE, abbr=FALSE)))

head(data.2)

```


**d)** Plot the data with the point color being determined by the day of the week variable.



```{r}

ggplot(data=data.2, aes(x=date,y=births,color=dow) )+
  geom_point()

```

