---
title: "assignment15"
author: "Noah Plant"
date: "2024-11-09"
output: pdf_document
---


### Exercise 2 {-}

Using the `datasets::trees` data, complete the following. This question refreshes create a linear model, graphing the linear model, and introduces using some LaTeX expressions on the graph.

**a)** Create a regression model for $y=$ `Volume` as a function of $x=$ `Height`.

```{r}


library(dplyr)
library(ggplot2)
library(broom)


myData<-datasets::trees

myData.model<-lm(Volume ~ Height ,data=myData)





```


**b)** Display the `summary` of the  model to view the y-intercept and slope of the regression line.

```{r}

summary(myData.model)

```

**c)** Using `ggplot2`, create a scatter plot of Volume vs Height. 

```{r}

plot.c<-myData%>%ggplot(aes(y=Volume,x=Height))+geom_point()

plot.c


```


**d)** Color the scatter using the `Girth` variable.

```{r}

plot.d<-myData%>%ggplot(aes(y=Volume,x=Height))+geom_point(aes(color=Girth))
plot.d

```


**e)** Modify the color scheme using a RColorBrewer palette.

```{r}

plot.e<-plot.d+scale_color_distiller(palette='Spectral')

plot.e

```


**f)** Create a nice white filled rectangle to add text information. The following might be useful.

```{r}

plot.f<-plot.e+annotate('rect', xmin=50, xmax=68, ymin=58, ymax=68,
         fill='white', color='black')


plot.f
```

**g)** Use the `broom` package to extract the coefficients of the best-fit line. Add this information as an annotation to the graph, which should follow a form that looks like $\hat{y}_i = (INTERCEPT) + (SLOPE)* x_i$. Place the annotation within the white text box.

```{r}

library(latex2exp)
cleanModel<-broom::tidy(myData.model)

intercept<-cleanModel[1,2]
slope<-cleanModel[2,2]


plot.g<-plot.f+annotate("text",x=59,y=61,label=latex2exp::TeX('$\\hat{y}_i = (-87.12361	) + (1.54335	)* x_i$')  )


plot.g

```


**h)** Use the `broom` package to extract the coefficient of determination $r^2$ from the model. Add the annotation to your graph, which should look something like $R^2 = (VALUE)$

```{r}

rsquared<-broom::glance(myData.model)

rsquared<-rsquared[1]

plot.h<-plot.g+annotate('text',x=59,y=65,label=latex2exp::TeX('$R^2 = (0.3579026	)$'))

plot.h

```


**i)** Add the regression line in red. There are several ways to do this.

```{r}

line<-broom::augment(myData.model, trees)

line

plot.i<-line%>%ggplot(aes(y=Volume,x=Height))+
  geom_point(aes(color=Girth))+
  geom_line(aes(y=.fitted))+
  scale_color_distiller(palette='Spectral')+
  annotate('rect', xmin=50, xmax=68, ymin=58, ymax=68,
         fill='white', color='black')+
  annotate("text",x=59,y=61,
           label=latex2exp::TeX('$\\hat{y}_i = (-87.12361	) + (1.54335	)* x_i$'))+
  annotate('text',x=59,y=65,label=latex2exp::TeX('$R^2 = (0.3579026	)$'))
  

  
  
plot.i


```


**j)** Properly label the axes of the graph.

```{r}

plot.j<-plot.i+ labs(x= "Height (ft)",y="Volume (ft^3)",color="Girth (in)")+theme(
    # Change x-axis font type
    axis.text.x = element_text(size = 12, face = "bold", family = "serif"),
    # Change y-axis font type
    axis.text.y = element_text(size = 12, face = "bold", family = "serif"),
    # Change x-axis title font type
    axis.title.x = element_text(size = 14, face = "bold"),
    # Change y-axis title font type
    axis.title.y = element_text(size = 14, face = "bold")
  )
  
plot.j
  



```


**k)** Add a descriptive title to the graph.

```{r}

plot.k<-plot.j+labs(title="Black Cherry Tree Dimensions")

plot.k

```

